---
title: "Function Development"
author: "[Skyler Lewis](mailto:slewis@flowwest.com)"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: true
    toc_depth: 3
    number_sections: false
    math_method:
      engine: webtex
      url: https://latex.codecogs.com/svg.image?
---

```{r}
library(tidyverse)
library(patchwork)
```

Function to pull flow-to-suitable-area curve for a particular comid/mainstem/watershed

```{r eval=FALSE, include=FALSE}
habitat_fsa <- function(reach, mainstem, watershed, 
                        habitat_type = "rearing", units = "ft") {
  
  mode_geom <- 
    if (!missing(reach)) "reach" else 
    if (!missing(mainstem)) "mainstem" else
    if (!missing(watershed)) "watershed"
  
  mode_unit <-
    if (units %in% c("ft", "feet", "ft2/lf")) "ft" else 
    if(units %in% c("ac", "acres")) "ac"

    fsa <- 
    switch(mode_geom,
           "reach" = 
             switch(mode_unit,
                    "ft" = 
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(comid == reach) |>
                      select(flow_cfs, 
                             habitat = wua_per_lf_pred),
                    "ac" =
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(comid == reach) |>
                      transmute(flow_cfs, 
                                habitat = wua_per_lf_pred * reach_length_ft / 43560)),
           "mainstem" = 
             switch(mode_unit,
                    "ft" = 
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(river_cvpia == mainstem) |>
                      mutate(habitat_ft2 = wua_per_lf_pred * reach_length_ft) |>
                      group_by(flow_cfs) |>
                      summarize(habitat = sum(habitat_ft2, na.rm = T) / 43560, 
                                .groups = "drop"),
                    "ac" =
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(river_cvpia == mainstem) |>
                      mutate(habitat_ft2 = wua_per_lf_pred * reach_length_ft) |>
                      group_by(flow_cfs) |>
                      summarize(habitat = sum(habitat_ft2, na.rm = T) / sum(reach_length_ft), 
                                .groups = "drop")),
           "watershed" = 
             switch(mode_unit,
                    "ft" = 
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(watershed_level_3 == watershed) |>
                      mutate(habitat_ft2 = wua_per_lf_pred * reach_length_ft) |>
                      group_by(flow_cfs) |>
                      summarize(habitat = sum(habitat_ft2, na.rm = T) / 43560, 
                                .groups = "drop"),
                    "ac" =
                      habistat::wua_predicted |>
                      filter(habitat == habitat_type) |>
                      filter(watershed_level_3 == watershed) |>
                      mutate(habitat_ft2 = wua_per_lf_pred * reach_length_ft) |>
                      group_by(flow_cfs) |>
                      summarize(habitat = sum(habitat_ft2, na.rm = T) / sum(reach_length_ft), 
                                .groups = "drop")))

    return(fsa)
    
}

# example mainstem
habitat_fsa(mainstem = "Feather River")

# example single comid
habitat_fsa(reach = 7978069)
```

```{r}
habitat_fsa_reach <- function(reach, habitat_type = "rearing", units = "ft") {
  
  mode_unit <-
    if (units %in% c("ft", "feet", "ft2/lf")) "ft" else 
    if(units %in% c("ac", "acres")) "ac"

  fsa <- 
    switch(mode_unit,
          "ft" = 
            habistat::wua_predicted |>
            filter(habitat == habitat_type) |>
            filter(comid == reach) |>
            select(flow_cfs, 
                   habitat = wua_per_lf_pred),
          "ac" =
            habistat::wua_predicted |>
            filter(habitat == habitat_type) |>
            filter(comid == reach) |>
            transmute(flow_cfs, 
                      habitat = wua_per_lf_pred * reach_length_ft / 43560))

    return(fsa)
    
}

habitat_fsa_reach(reach = 7978069) |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = habitat)) +
  scale_x_log10()
```


```{r}
habitat_fsa_reach_scaled <- function(reach, multiplier, ...) {
  
  fsa <- habitat_fsa_reach(reach, ...)
  
  return(tibble(flow_cfs = fsa$flow_cfs,
                habitat = approx(x = fsa$flow_cfs * multiplier,
                                 y = fsa$habitat,
                                 xout = fsa$flow_cfs,
                                 rule = 2:2,
                                 method = "linear",
                                 na.rm = F)$y))
    
}

habitat_fsa_reach_scaled(reach = 7978069, multiplier = 0.5) |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = habitat)) +
  scale_x_log10()
```

```{r}
habitat_fsa <- function(reach, mainstem, watershed, 
                        habitat_type = "rearing", units = "ft") {
  
  mode_geom <- 
    if (!missing(reach)) "reach" else 
    if (!missing(mainstem)) "mainstem" else
    if (!missing(watershed)) "watershed"
  
  mode_unit <-
    if (units %in% c("ft", "feet", "ft2/lf")) "ft" else 
    if(units %in% c("ac", "acres")) "ac"
  
  if (mode_geom == "reach") {
    
    habitat_fsa_reach(reach = reach,
                      habitat_type = habitat_type,
                      units = units)
    
  } else if (mode_geom %in% c("mainstem", "watershed")) {
    
    group_var <- 
      switch(mode_geom,
             "mainstem" = "river_cvpia",
             "watershed" = "watershed_level_3")
        
    flow_xw <- 
      switch(mode_geom,
             "mainstem" = habistat::cv_mainstems_flow_xw,
             "watershed" = habistat::cv_watersheds_flow_xw)
    
    comids <-
      switch(mode_geom,
             "mainstem" = 
               habistat::cv_mainstems |>
               filter(river_cvpia == mainstem) |>
               pull(comid),
             "watershed" = 
               habistat::cv_watersheds |>
               filter(river_cvpia == mainstem) |>
               pull(comid))
    
    scaled_predictions <-
      flow_xw |>
      filter(comid %in% comids) |>
      mutate(fsa = map2(comid, multiplier, function(x, y) {
        habitat_fsa_reach_scaled(reach = x,
                                 multiplier = y,
                                 habitat_type = habitat_type,
                                 units = "ft")})) |> # output in units ft2/ft
      inner_join(habistat::flowline_attr |> select(comid, reach_length_ft),
                 by = join_by(comid)) |>
      unnest(fsa) |>
      mutate(habitat_ft2 = habitat * reach_length_ft)
    
    switch(mode_unit,
           "ft" = 
             scaled_predictions |>
             group_by(flow_cfs) |>
             summarize(habitat = sum(habitat_ft2, na.rm = T) / sum(reach_length_ft), 
                       .groups = "drop"),
           "ac" = 
             scaled_predictions |>
             group_by(flow_cfs) |>
             summarize(habitat = sum(habitat_ft2, na.rm = T) / 43560, 
                       .groups = "drop"))

  }
  
}

habitat_fsa(mainstem = "Feather River", units = "ac") |>
  ggplot() + 
  geom_line(aes(x = flow_cfs, y = habitat)) +
  scale_x_log10()
```


Function to pull duration HSI by flow

```{r}
habitat_drc <- function(streamgage,
                        habitat_type = "rearing",
                        run = "fall",
                        wy_group = "Dry",
                        gradient = "Valley Foothill") {
  
  drc_filtered <-
    habistat::streamgage_duration_rating_curves |>
    filter((station_id == str_to_lower(streamgage)) &
             (run == str_to_lower(run)) &
             (habitat_type == str_to_lower(habitat_type)) &
             (wy_group == str_to_title(wy_group)))
  
  varname <- case_when(
    habitat_type == "spawning" ~ "durhsi_spawning",
    str_to_title(gradient) == "Valley Lowland" ~ "durhsi_rearing_vl",
    TRUE ~ "durhsi_rearing_vf")
  
  drc_selected <-
    drc_filtered$data[[1]] |>
    select(flow_cfs = model_q, 
           durhsi = !!sym(varname))
  
  return(drc_selected)
  
}

habitat_drc("fsb", habitat_type="rearing", gradient = "Valley Foothill") |>
  ggplot() + geom_line(aes(x = flow_cfs, y = durhsi))
```
  

Function to pull duration HSI by flow, weighted for a comid based on DA and MAP ratios

```{r}
habitat_drc_weighted <- function(streamgage,
                                 reach,
                                 scale = "durhsi",
                                 ...) {
  
  gage_attr <-
    habistat::streamgage_da_attr |>
    filter(station_id == str_to_lower(streamgage)) |>
    as.list()
  
  comid_attr <-
    habistat::flowline_attr |>
    filter(comid == reach) |>
    select(comid, da_area_sq_km, da_ppt_mean_mm) |>
    as.list()
  
  multiplier <-
    (comid_attr$da_area_sq_km / gage_attr$da_gage) *
    (comid_attr$da_ppt_mean_mm / gage_attr$pc_gage)
  
  drc_unweighted <-
    habitat_drc(streamgage, ...)
  
  result <-
    switch(scale,
           "durhsi" = 
             tibble(flow_cfs = drc_unweighted$flow_cfs,
                    durhsi = approx(x = drc_unweighted$flow_cfs * multiplier, 
                                    xout = drc_unweighted$flow_cfs, 
                                    y = drc_unweighted$durhsi,
                                    method = "linear",
                                    rule = 2:2,
                                    na.rm = FALSE)$y),
           "flow" = 
             drc_unweighted |>
             mutate(flow_cfs = flow_cfs * multiplier))
  return(result)
  
}

habitat_drc_weighted("FSB", habitat_type="rearing", gradient = "Valley Foothill", reach = 7978069, scale="durhsi")  |>
  ggplot() + geom_line(aes(x = flow_cfs, y = durhsi))
```

Function to pull FSA curve - weighted by duration - for single reach

```{r}
habitat_fsa_duration_reach <- function(reach,
                                       streamgage,
                                       habitat_type = "rearing", 
                                       units = "ft",
                                       run = "fall",
                                       wy_group = "Dry",
                                       gradient = "Valley Foothill") {
  
  fsa <- 
    habitat_fsa(reach = reach,
                habitat_type = habitat_type,
                units = units) 
  
  drc <-
    habitat_drc_weighted(streamgage = streamgage,
                         reach = reach,
                         habitat_type = habitat_type,
                         run = run,
                         wy_group = wy_group,
                         gradient = gradient) 
  
  fsa_weighted <-
    habistat::duration_apply_dhsi_to_fsa_curve(fsa = fsa,
                                               drc = drc, 
                                               # var names
                                               fsa_q = flow_cfs,
                                               fsa_wua = habitat,
                                               drc_q = flow_cfs,
                                               drc_dhsi = durhsi)
  
  return(tibble(flow_cfs = fsa$flow_cfs,
                habitat = approx(x = fsa_weighted$q,
                                 xout = fsa$flow_cfs,
                                 y = fsa_weighted$durwua,
                                 rule = 2:2, 
                                 na.rm = F)$y))
  
}

# example non-temporal
habitat_fsa(reach = 7978069) 

# example temporal
habitat_fsa_duration_reach(reach = 7978069, streamgage = "FSB")
```

Function to pull FSA curve - weighted by duration - with watershed aggregation logic

```{r}
habitat_fsa_duration <- function(reach,
                                 mainstem,
                                 watershed,
                                 streamgage,
                                 habitat_type = "rearing", 
                                 units = "ft",
                                 run = "fall",
                                 wy_group = "Dry",
                                 gradient = "Valley Foothill") {
  
  mode_geom <- 
    if (!missing(reach)) "reach" else 
    if (!missing(mainstem)) "mainstem" else
    if (!missing(watershed)) "watershed"
  
  if (mode_geom == "reach") {
    
    habitat_fsa_duration_reach(reach = reach,
                               streamgage = streamgage,
                               habitat_type = habitat_type, 
                               units = units,
                               run = run,
                               wy_group = wy_group,
                               gradient = gradient)
    
  } else if (mode_geom %in% c("mainstem", "watershed")) {
    
    group_var <- 
      switch(mode_geom,
             "mainstem" = "river_cvpia",
             "watershed" = "watershed_level_3")
        
    flow_xw <- 
      switch(mode_geom,
             "mainstem" = habistat::cv_mainstems_flow_xw,
             "watershed" = habistat::cv_watersheds_flow_xw)
    
    comids <-
      switch(mode_geom,
             "mainstem" = 
               habistat::cv_mainstems |>
               filter(river_cvpia == mainstem) |>
               pull(comid),
             "watershed" = 
               habistat::cv_watersheds |>
               filter(river_cvpia == mainstem) |>
               pull(comid))
    
    scaled_predictions <-
      flow_xw |>
      filter(comid %in% comids) |>
      mutate(fsa = map2(comid, multiplier, function(x, y) {
        habitat_fsa_reach_scaled(reach = x,
                                 multiplier = y,
                                 habitat_type = habitat_type,
                                 units = "ft")})) |> # output in units ft2/ft
      mutate(drc = map2(comid, streamgage, function(x, y) {
        habitat_drc_weighted(reach = x,
                             streamgage = y,
                             run = run,
                             wy_group = wy_group,
                             gradient = gradient,
                             scale = "flow")})) |>
      mutate(fsa_x_drc = map2(fsa, drc, function(f, d) {
        fsa_weighted <- 
          habistat::duration_apply_dhsi_to_fsa_curve(fsa = f,
                                                     drc = d, 
                                                     # var names
                                                     fsa_q = flow_cfs,
                                                     fsa_wua = habitat,
                                                     drc_q = flow_cfs,
                                                     drc_dhsi = durhsi)
        return(tibble(flow_cfs = f$flow_cfs,
                      habitat = approx(x = fsa_weighted$q,
                                       xout = f$flow_cfs,
                                       y = fsa_weighted$durwua,
                                       rule = 2:2, 
                                       na.rm = F)$y))
        })) |>
      inner_join(habistat::flowline_attr |> select(comid, reach_length_ft),
                 by = join_by(comid)) |>
      select(-fsa, -drc) |>
      unnest(fsa_x_drc) |>
      mutate(habitat_ft2 = habitat * reach_length_ft)
    
    switch(mode_unit,
           "ft" = 
             scaled_predictions |>
             group_by(flow_cfs) |>
             summarize(habitat = sum(habitat_ft2, na.rm = T) / sum(reach_length_ft), 
                       .groups = "drop"),
           "ac" = 
             scaled_predictions |>
             group_by(flow_cfs) |>
             summarize(habitat = sum(habitat_ft2, na.rm = T) / 43560, 
                       .groups = "drop"))

  }
  
}

habitat_fsa_duration(reach = 7978069, streamgage = "FSB")

habitat_fsa(mainstem = "Feather River")
habitat_fsa_duration(mainstem = "Feather River", streamgage = "FSB")
```





Function to interpolate a FSA curve TS weighted by duration

Function to extract suitable area at a given flow or vector of flows

```{r}
habitat_predict <- function(x, 
                            reach, mainstem, watershed, 
                            streamgage,
                            fsa_x,
                            fsa_y,
                            habitat_type = "rearing", 
                            units = "ft",
                            run = "fall",
                            wy_group = "Dry",
                            gradient = "Valley Foothill") {
  
  mode_geom <- 
    if (!missing(reach)) "reach" else 
    if (!missing(mainstem)) "mainstem" else
    if (!missing(watershed)) "watershed"
  
  mode_unit <-
    if (units %in% c("ft", "feet", "ft2/lf")) "ft" else 
    if(units %in% c("ac", "acres")) "ac"
  
  if (!missing(streamgage)) {
    
    fsa <- habitat_fsa_duration(reach = reach,
                                mainstem = mainstem,
                                watershed = watershed,
                                streamgage = streamgage,
                                habitat_type = habitat_type,
                                units = units,
                                run = run,
                                wy_group = wy_group,
                                gradient = gradient)
    
  } else if (isTRUE((!missing(fsa_x) && (!missing(fsa_y))))) {
    
    fsa <- tibble(flow_cfs = fsa_x,
                  habitat = fsa_y)
    
  } else {
    
    fsa <- habitat_fsa(reach = reach,
                       mainstem = mainstem,
                       watershed = watershed,
                       habitat_type = habitat_type,
                       units = units)
    
  }

  result <- approx(x = if_else(fsa$flow_cfs > 0, log10(fsa$flow_cfs), 0), 
                   y = fsa$habitat, 
                   xout = if_else(x > 0, log10(x), 0),
                   method = "linear",
                   rule = 2:2)
  
  return(result$y)
  
}

# example non-temporal prediction given comid
habitat_predict(c(0, 10, 150, 300, 9999), reach = 7978069)

# example temporal prediction given comid and streamgage
habitat_predict(c(0, 10, 150, 300, 9999), reach = 7978069, streamgage = "FSB")

# example non-temporal prediction given mainstem
habitat_predict(c(0, 10, 150, 300, 9999), mainstem = "Feather River")

# example temporal prediction given comid and streamgage
habitat_predict(c(0, 10, 150, 300, 9999), mainstem = "Feather River", streamgage = "FSB")

# example manual prediction given custom provided flow-to-suitable-area curve
habitat_predict(c(0, 10, 150, 300, 9999), fsa_x = c(0, 100, 1000), fsa_y = c(47, 470, 4700))
```













Example application

```{r}
# random sample time sweries
tsdata <- 
  tibble(index = seq(0, 364, 1)) |>
  mutate(obs_date = ymd("2046-10-01") + index) |>
  mutate(delta = rnorm(n(), 0, 100),
         flow_cfs = 1000,
         flow_cfs = lag(flow_cfs) + cumsum(delta))

plt_ts <- 
  tsdata |>
  ggplot() + geom_line(aes(x = obs_date, y = flow_cfs)) +
  scale_y_log10(limits = c(50, 15000))

plt_fsa_transposed <-
  habitat_fsa(mainstem = "Feather River") |>
  ggplot() + geom_line(aes(x = flow_cfs, y = habitat))  +
  scale_x_log10(limits = c(50, 15000)) +
  scale_y_continuous(limits = c(0, 1000))

plt_fsa <-
  habitat_fsa(mainstem = "Feather River") |>
  ggplot() + geom_line(aes(y = flow_cfs, x = habitat))  +
  scale_y_log10(limits = c(50, 15000)) +
  scale_x_continuous(limits = c(0, 1000))

plt_hab <- 
  tsdata |>
  mutate(habitat = habitat_predict(flow_cfs, mainstem = "Feather River")) |>
  ggplot() + geom_line(aes( x = obs_date, y = habitat))

(plt_fsa / plt_fsa_transposed) | (plt_ts / plt_hab) + plot_layout(axes = "collect")
```














